Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-05-28 edubots_surveys.Rnw
Update State of the Art #6

It is essential, when implementing a educational chatbot strategy, to asses the needs of the community prior to implementation \cite{tsivitanidouusers}.

% TODO: state of the art

% student/teacher preferences towards messaging and chatbots (#6)

Conversational programs to facilitate communication between computers and humans using natural language have existed for many years \cite{Thompson2007,SHAH2016}. 
However, those were very limited systems in terms of the capacity of the models used at the time \cite{Hutchens1996}.

In recent years, several authors have carried out studies on the convenience of integrating chatbots in teaching, evaluating their characteristics. 
Thus, Coniam \cite{Coniam2014} evaluates several systems, concluding that they are still in an early stage to be used with guarantees.
Likewise, Tallyn et al. \cite{Tallyn2018} evaluates a system related to a very specific topic that is suitable for answering specific questions using a chat. Finally, Smutny et al. \cite{Smutny2020} present a review study examining the educational chatbots in Facebook Messenger with focus to identify characteristics such as language, subject matter and platform.

According to \cite{Roblyer2010}, using either tools such as mobile devices or teaching strategies based on gamification \cite{Yildirim2017} can improve student motivation. In this sense, Pimmer et al. \cite{Pimmer2019} carried out a series of experiments with technological tools whose results showed positive perception and acceptance of the use of technology in teaching and learning, since students are used to using those media with which they can express themselves freely.

% Categorization and examples of what they can do, adapted to questions in the survey   
A possible categorization of chatbots in Education splits them according to their purpose into service- and teaching-oriented \cite{perez2020rediscovering}. The first class includes chatbots that provide service support used e.g. to solve questions about the enrollment \cite{santoso2018dinus}, registration, or academic requirements for exchange students \cite{ranoliya2017chatbot}. Focusing on the use at class, service-oriented chatbots would help to introduce students to each other (as done with LISA \cite{dibitonto2018chatbot}), would provide information about schedules and deadlines, recommend general learning materials for the next classes, or individually inform about grades. Teaching-oriented chatbots are more sophisticated chatbots that set goals and monitor learning progress. A simple example is described in \cite{coronado2018cognitive} where authors proposed agents that store learning materials to be provided on demand to students. Also, more complex tutoring systems \cite{crockett2017predicting} individually analyze the learning state of students to provide the next most suitable material in their learning path.

% Point: chatbots are accepted if they help to achieve learning goals
Regarding the factors for the adoption of chatbots in Education, many studies have focused on the evaluation of technology acceptance or usability such as \cite{Pimmer2019, Roblyer2010}. However, Higher Education is a special domain where according to \cite{hobert2019you}, specific pedagogical factors such as learning success or increased motivation are more important.

>>>>

The Use of Mobile Instant Messaging (MIM) Platforms in Higher Education
The widespread and rapid adoption of free MIM tools/platforms such as WhatsApp, Telegram, WeChat and Facebook Messenger stems from their simplicity, ease of use and multi-modality (i.e. video, audio, text) (Tang and Hew, 2017).  Using these tools in higher education can facilitate the delivery of personalised learning that occurs anytime anywhere, promote collaborative learning experiences and group discussions (Panah and Baber, 2020). WhatsApp is the most popular MIM platform used by educators to give assignments’ feedback to students, support course discussions and provide learning resources in informal learning settings (Panah and Baber, 2020).  The use of WhatsApp in higher education could enhance social presence (Tang and Hew, 2017) and foster trust relationships between educators and students embedded in the social learning process (Gachago, Strydom, Hanekom, Simons and Walters, 2015).

Nevertheless, there are challenges when using MIM tools that occur due to the blurring of boundaries between academic and private life. This can lead to technostress (Gachago et al., 2015), difficulty in managing responsibilities especially among mature students and lack of privacy (Tang and Hew, 2017).  Students’ dropout of the MIM groups, as they can leave groups at any time can hinder their learning and undermine educators’ efforts (Mwakapina, Mhandeni, and Nyinondi, 2016).  In addition, there is a need to set rules and norms for these MIM groups in order to maintain the safety of these online communities for students (Abbas et al., 2021b) without affecting students’ ownership and control that is vital to advance their learning (Gachago et al., 2015).

The Use of Chatbots in Higher Education
The use of conversational agents or chatbots in higher education is still at its infancy (Yang and Evans, 2019). Nevertheless, recent studies examining their positive impact on students’ academic performance (Pérez, Daradoumis and Puig, 2020) and engagement (Studente, Ellis and Garivaldis, 2020; Abbas et al., 2021b) have led to a growing interest in using this technology in classes. Indeed, the use of chatbots in collecting course feedback from students in higher education improved students’ response quality and boosted their enjoyment levels (Abbas, Pickard, Atwell and Walker, 2021a).  According to Roblyer et al. (2010), using either tools such as mobile devices or teaching strategies based on gamification (Yildirim, 2017) can improve student motivation. In this sense, Pimmer et al.  (2019) carried out a series of experiments with technological tools whose results showed positive perception and acceptance of the use of technology in teaching and learning.

Several higher education chatbots’ evaluation studies were undertaken. For instance, an evaluation review study presented by Smutny and Schreiberova (2020) examined 47 educational chatbots implemented in Facebook Messenger with the focus to identify characteristics and quality metric such as language, subject matter and platform. Another study aimed to categorise educational chatbots according to their purpose into service-oriented and teaching-oriented (Pérez, Daradoumis and Puig, 2020). The first category includes those that provide service support such as the chatbot Ask Holly (University of Durham, 2021) and Dina (Santoso et al., 2018), both chatbots respond to students’ questions about enrolment and registration. Ask L.U. (Lancaster University, 2019) answers students’ frequently asked questions about timetables, grades, tutors and societies. LISA (Dibitonto et al., 2018) and Differ (Differ, 2021) facilitate breaking the ice between new students by introducing them to each other.  Ranoliya et al. (2017) proposed a generic chatbot for university students that is able to answer their frequently asked questions.

On the other hand, teaching-oriented chatbots are more sophisticated as they set personalised learning outcomes and monitor learning progress. For instance, Fernoagă, Stelea, Gavrilă, and Sandu (2018) reported on ‘eduAssistant’ a virtual teaching assistant chatobt developed on the Telegram messaging platform. In this study, the Telegram platform was chosen because it is easy to use, students are familiar with its features and it enables students to exchange messages in different formats (text, audio and video) (Fernoagă et al., 2018). In addition, Telegram could operate on all devices and operating systems. The ‘eduAssistant’ chatbot acts as an automatic agent between the teacher-content-student, facilitating real-time feedback loops and providing a personalised learning experience relevant to students’ skills and knowledge. Using this chatbot, educators can create interactive instances in their lectures where they pose questions to their students and the chatbot assists students who need further help by giving those more hints and reporting to their educator’s dashboard (Fernoagă et al., 2018). This can help educators locate those students that need their attention and send them more educational resources relevant to their academic attainment.  
Furthermore, Sjöström et al. (2018) proposed a conceptual architecture for teaching-oriented chatbots in higher education.  This conceptual architecture is based on a systematic literature review of previous studies examining the design of chatbots in higher education and a content analysis of student emails and discussion forum posts of four instances of a Java programming course.  The study outlined several design considerations, among them, the authors emphasised the importance of developing chatbots in platforms that students and educators are familiar with and can easily access (i.e. Facebook Messenger) which confirms with Hobert’s (2019) and Fernoagă, Stelea, Gavrilă, and Sandu’s (2018)  studies. In addition, Sjöström et al. argued that a conceptualision of learners’ questions could aid designers in integrating the appropriate types of questions that the chatbots should support for different courses.  Coronado et al. (2018) proposed agents that store learning materials to be provided on demand to students. Crockett et al. (2017) reported on tutoring systems, which can perform initial assessments of students’ understanding and provid learning material that would advance their understanding to the next level.
Regarding the factors for the adoption of chatbots in higher education, many studies have focused on the evaluation of technology acceptance and usability (Roblyer et al., 2010; Pimmer et al., 2019). However, higher education is a special domain where according to Hobert (2019), specific pedagogical factors such as learning success or increased motivation are more important.  Therefore, to develop efficient chatbots for higher education, all stakeholders’ (i.e. educators, students, institutions, etc) expectations needs to be carefully collected and taken into consideration (Sjöström et al., 2018; Tsivitanidou and Ioannou, 2020).
